A Classical Density-Functional Theory for Describing Water Interfaces

### 2012-05-05 papers/water-SAFT/paper.tex

A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
describes water as non-interacting molecules in a an
effective external potential.  This functional contains the exact
entropy for no-interacting molecules and, among other things, can 
predict the nonlinear dielectric response in water but must be paramaterized
for different temperatures.

\needsworknow{Add smooth transition from others' work to ours}

\section{Theory and Methods}
We use classical DFT to construct a new functional that will predict
the density of water molecules in a classical system.  The model we use
starts with a treatment of each individual water molecule as a purely
classical hard sphere and adds association terms based on Statistical 
Associating Fluid Theory (SAFT). Water is unlike simple
fluids (which are mainly defined by van der Waals forces and weak electrostatic
interations) because it requires direct treatment of association effects. The strong 
hydrogen bonds and polar
interactions in water are ideal for describing with SAFT, which was designed for
use with hydrogen-bonding fluids. It is based on Wertheim's first-order
thermodynamic perturbation theory (TPT1)
\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
wertheim1986fluidsIV} which provides a useful way to account for strong
associative interations between molecules. SAFT can account for the association energies
as well as any effects due to clusters or chains forming \fixme{not true}
in the liquid\cite{muller2001molecular}.  \needsworknow{Expand this a
  bit to accurately describe SAFT and its advantages.}

@@

We simulate four hydrophobic rods in water using the same radii as our
two rods ($r = 0.3$~nm~-~$1.0$~nm) and
separation distances between $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
the rods in a 'square' arrangement and observed a similar behavior as the two rods
where vapor is present in the space between the rods until a critical separation
where it becomes filled with water and the rods behave like separate single rods
(shown in Fig. \ref{fig:density-four-rods}).  Free energy per length
is plotted against rod separation in Figure
\ref{fig:four-rods-energy-vs-distance} with a slight disagreement of
the predicted$2\gamma$ force per unit length.

The 'rounded square' formed by the area of the four rods and the space
inbetween them has a circumference estimated by
\begin{equation}
C_{rs} = 2 \pi r + 8r + 4d.
\end{equation}
We again equate this circumference with the total circumference of
only the four rods, which results in a critical separation of
\begin{equation}
d_{c4} = \frac{3 \pi - 4}{2} r. \label{criticalfourrods}
\end{equation}
We see in Figure \ref{fig:four-rods-energy-vs-distance} that the $r = 0.3$~nm
rod has the largest difference between simulation and prediction with
a 23\% difference, and better agreement for larger radii rods (between 0.5\% and 5.6\%).

>>>>

A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
describe water as non-interacting molecules in a an
effective external potential.  This functional contains the exact
entropy for no-interacting molecules and, among other things, can 
predict the nonlinear dielectric response in water, but it must be paramaterized
for different temperatures.

\needsworknow{Add smooth transition from others' work to ours}
Using classical DFT, we aim to find an equation of state for water
that describes interactions qualitatively and gives accurate
quantitative results within a reasonable temperature range without
needing extensive paramaterization.

\section{Theory and Methods}
We use classical DFT to construct a new functional that will predict
the density of water molecules in a classical system.  The model we use
starts with a treatment of each individual water molecule as a purely
classical hard sphere and adds association terms based on Statistical 
Associating Fluid Theory (SAFT). Water is unlike simple
fluids (which are mainly defined by van der Waals forces and weak electrostatic
interations) because it requires direct treatment of association effects. The strong 
hydrogen bonds and polar
interactions in water are ideal for describing with SAFT, which was designed for
use with hydrogen-bonding fluids. It is based on Wertheim's first-order
thermodynamic perturbation theory (TPT1)
\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
wertheim1986fluidsIV} which provides a useful way to account for strong
associative interations between molecules. SAFT can account for the association energies
as well as any effects due to clusters or chains forming \fixme{not true}
in the liquid\cite{muller2001molecular}. SAFT assumes a repulsive hard
sphere core for each atom with a number of short-ranged off center
attractive sites. This model applied to give equation of state for
various systmes and can be applied to inhomogeneous fluids.
\needsworknow{Expand this a
  bit to accurately describe SAFT and its advantages.}

@@

We simulate four hydrophobic rods in water using the same radii as our
two rods ($r = 0.3$~nm~-~$1.0$~nm) and
separation distances between $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
the rods in a 'square' arrangement and observed a similar behavior as the two rods
where vapor is present in the space between the rods until a critical separation
where it becomes filled with water and the rods behave like separate single rods
(shown in Fig. \ref{fig:density-four-rods}).  Free energy per length
is plotted against rod separation in Figure
\ref{fig:four-rods-energy-vs-distance} with a slight disagreement of
the predicted force per unit length.

The 'rounded square' formed by the area of the four rods and the space
between them has a circumference estimated by
\begin{equation}
C_{rs} = 2 \pi r + 8r + 4d.
\end{equation}
We again equate this circumference with the total circumference of
the four separate rods, which results in a critical separation of
\begin{equation}
d_{c4} = \frac{3 \pi - 4}{2} r. \label{criticalfourrods}
\end{equation}
We see in Figure \ref{fig:four-rods-energy-vs-distance} that the $r = 0.3$~nm
rod has the largest difference between simulation and prediction with
a 23\% difference, and better agreement for larger radii rods (between 0.5\% and 5.6\%).
